---
title: "Boechera_ZS_2.15.22"
author: "Zachary Snider"
date: "4/15/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggfortify)
library(corrplot)
library(factoextra)
library(dplyr)
library(vegan)
library(adegenet)
```

To look at the default color palette used
```{r}
library(scales)
show_col(hue_pal()(4))
```

Setting directory
```{r}
#setwd("~/Desktop/Boechera_GIS")
#setwd("C:/Users/Daniel/Desktop/BoecheraStress-main")
setwd("~/Documents/GitHub/BoecheraStress")
```

Loading the data
```{r}
bclim <- read.csv("boechera_climate_data.csv")
colnames(bclim) <- c('PopID','Individual','Region','Species','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans')
```

To subset to only b. californica, run this chunk
```{r}
bclim <- bclim %>%
  filter(Species=="californica")
```

Geographic distribution of Boechera Samples
```{r}
plot1 <- ggplot(bclim, aes(x=Latitude,y=Elevation,colour=Region)) + 
  geom_point(size=4) + 
  theme_bw() +
  theme(text = element_text(size=28)) +
  ylab("Elevation (m)")
plot1
```

Histogram of population counts
```{r}
plot5 <- ggplot(bclim, aes(x=PopID,fill=Region)) +
  geom_bar() +
  theme_bw() +
  theme(text = element_text(size=20)) +
  xlab("Populations")
plot5
plot6 <- ggplot(bclim, aes(x=Region, fill=Region)) +
  geom_bar(stat="count") +
  theme_bw() +
  theme(text = element_text(size=20)) +
  geom_text(stat='count', aes(label=..count..), vjust=-.5)
plot6
```

Plots against precipitation
```{r}
plot7 <-ggplot(bclim, aes(y=PPT, x=Region, fill=Region)) +
  geom_boxplot() +
  theme_bw() +
  ylab("Average Annual Precipitation (mm)") +
  theme(text = element_text(size=20))
  #+ stat_summary(geom = "crossbar", width=1, fatten=.8, color="black", fun.data = function(x){ return(c(y=median(x), ymin=median(x), ymax=median(x))) })
plot7
```
  
Plot against soil type (No longer working without soil info)
```{r, error=TRUE}
#plot8 <-ggplot(boechera, aes(x=Latitude,y=Elevation,colour=PTYPE)) + 
  #geom_point(size=4) + theme_bw() +
  #theme(text = element_text(size=28)) +
  #ylab("Elevation (m)")
#plot8
```

PRISM GIS data
```{r}
library(prism)
#get_prism_normals(type="tmean", resolution = "4km", annual = TRUE, keepZip= TRUE)
#get_prism_normals("ppt", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("tmean", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("tmin", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("tmax", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("tdmean", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("vpdmin", "800m", annual=TRUE, keepZip= TRUE)
#get_prism_normals("vpdmax", "800m", annual=TRUE, keepZip= TRUE)
library(raster)
library(rgdal)
#rasnam <- "PRISM_tmean_30yr_normal_800mM3_annual_bil.bil"
#ras<-raster(rasnam) # identify as raster
#prj<-toString(crs(ras)) #get raster projection
#plot(ras)
```

Mean annual temperatures
```{r}
plot9 <- ggplot(bclim, aes(y=tmean, x=Region, fill=Region)) +
  geom_boxplot() +
  theme_bw() +
  ylab("Mean Annual Temperature (°C)") +
  theme(text = element_text(size=20))
plot9
```

PCA 
```{r}
pca.boc = prcomp(bclim[,c(7,9:15,17,18)], center=TRUE, scale=TRUE)
summary(pca.boc)
screeplot(pca.boc, n = 11)
```

Autoplot with PCA loadings
```{r, dpi=300}
autoplot(pca.boc, data = bclim[,c(7,9:15,17,18)], 
		label = FALSE, loadings = TRUE, 
		loadings.colour = "black",
		loadings.label = TRUE,
		loadings.label.hjust = 1.2) +
  theme_bw() 
```

PCA biplot with convex hulls
```{r rawReducedBiPlot, dpi = 300}
fviz_pca_biplot(pca.boc, col.ind=bclim$Region, addEllipses=TRUE,
		mean.point=TRUE, ellipse.type = 'convex',
		geom.var=c('point','text'), repel=TRUE)
```

Reducing the data set by using the variables that contribute most to PCs 1 and 2
```{r}
var = get_pca_var(pca.boc)
fviz_contrib(pca.boc, choice='var', axes = 1:2)
f=fviz_contrib(pca.boc, choice = "var", sort.val=c("desc"), axes = 1:2)
dat=f$data
compounds=rownames(dat[dat$contrib>9,])
reduced.data=bclim[,compounds]
```

Creating new PCA based on reduced dataset
```{r}
pca.red = prcomp(reduced.data, center=TRUE, scale.=TRUE)
summary(pca.red)
```

Plotting the reduced PCA
```{r rawReducedBiPlot, dpi =1 300}
fviz_pca_biplot(pca.red, col.ind=bclim$Region, addEllipses=TRUE,
		mean.point=TRUE, ellipse.type = 'convex',
		geom.var=c('point','text'), repel=TRUE, title = "PCA Biplot using climate") +
  theme(text = element_text(size = 20),
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 20))
```


```{r}
find.clusters(bclim[7,9:15,17,18])
```




###
Analysis of Similarity
```{r}
library(vegan)
# x is a matrix of response variables, grouping is the independent variable
sim1 = anosim(x = as.matrix(bclim[,c(7,9:15,17,18)]), grouping = boechera$Pop.ID)
summary(sim1)
perm = permustats(sim1) # Extract the permutation results from sim1 
densityplot(perm) # plot the density of R, the test statistic for the anosim

```


##Windham data
```{r}
wind <- read.csv("arcuata_windham.csv")
wind2 <- read.csv("arcuata_CA.csv", header=TRUE)
arcuata<-merge(wind,wind2,by=c("Extrac"))
arcuata$I3 <- as.numeric(arcuata$I3)
library(dplyr)
arcuata <- arcuata %>% 
     mutate_at(c(8:42), as.numeric)
```

```{r}
pca.arc = prcomp(arcuata[,c(8,10,11,13,14,16,18,20,22,24,26,27,31,33,39)], center=TRUE, scale=TRUE, na.action(na.omit))
summary(pca.arc)
pca.arc2 = prcomp(arcuata[,c(8:42)], center=TRUE, scale=TRUE, na.action(na.omit))
summary(pca.arc2)
```











