---
title: "msat.popgen"
author: "Zachary Snider"
date: "2022-10-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

Analysis of Boechera Microsats of SoCal Populations
#In Github repository BoecheraStress

Loading packages
```{r setup, include=FALSE}
library(ggplot2)
library(ggfortify)
library(tidyverse)
library(polysat)
library(poppr)
library(pegas)
library(adegenet)
library(reshape2)
library(vegan)
library(lattice)
library(RColorBrewer)
library(mmod)
library(ape)
library(magrittr)
library(ggpubr)
library(dartR)
```

```{r}
#setwd("~/Desktop/microsats/Rcode")
setwd("~/Documents/GitHub/BoecheraStress")
```

PolySat Analysis
```{r}
msbc <- read.GeneMapper("polysat_msats_californica1")
Description(msbc) <- "Boechera californica Microsats collected by ZS"
#Name populations
PopNames(msbc) <- c("SD", "LA", "SJ","SB") 
gpops <- c(1,1,1,1,1,1,1,1,1,1,
           1,1,1,1,1,1,1,1,1,1,
           1,1,1,1,1,1,1,1,1,1,
           1,1,2,2,2,2,2,2,2,2,
           2,2,3,3,3,3,3,3,3,3,
           3,3,3,1,1)
PopInfo(msbc) <- gpops
PopInfo(msbc)
#Assign ploidy
Ploidies(msbc) <- c(3)
Ploidies(msbc)
#Checking populations
Samples(msbc, populations = "SD")
Samples(msbc, populations = "LA")
Samples(msbc, populations = "SJ")
Samples(msbc, populations = "SB")

#Assigning length of microsats
Usatnts(msbc) <- c(25, 17, 20, 21, 19, 20, 20, 20, 20, 19, 20, 19, 20, 18, 20)
Usatnts(msbc)
```

PCA plots
```{r}
##Genetic distances between individuals
testmatrix <- meandistance.matrix(msbc,distmetric=Bruvo.distance)
pca1 <- cmdscale(testmatrix)
mycol <- c("#999999","#E69F00","#56B4E9","#009E73")
ggplot(pca1, aes(x=pca1[,1], y=pca1[,2], color= as.character(PopInfo(msbc)))) +
  geom_point(size=2) +
  
  stat_ellipse(geom = "polygon",
               aes(fill = as.character(PopInfo(msbc))), 
               alpha = 0.2) +
  labs(x="Principal Component 1", y="Principal Component 2") +
  labs(title = "PCA with Bruvo Distance using only B. californica") +
  scale_fill_discrete(name ="Regions", labels = c("SD","LA","SJ")) +
  guides(color = "none") + 
  theme_bw(base_size=22)
#Export with 1100 x 700 pixels
#Bruvo distance = takes mutational distance between alleles into account (2004)
citation("polysat")
testmatrix2 <- meandistance.matrix(msbc, distmetric=Lynch.distance)
pca2 <- cmdscale(testmatrix2)
ggplot(pca2, aes(x=pca2[,1], y=pca2[,2], color= as.character(PopInfo(msbc)))) +
  geom_point(size=2) +
  stat_ellipse(geom = "polygon",
               aes(fill = as.character(PopInfo(msbc))), 
               alpha = 0.2) +
  labs(x="Principal Component 1", y="Principal Component 2") +
  labs(title = "PCA with Lynch Distance using only B. californica") +
  scale_fill_discrete(name ="Regions", labels = c("SD","LA","SJ","SB")) +
  guides(color = "none") +
  theme_bw(base_size=22)
#Export with 1100 x 700 pixels
#Lynch distance = a simple band-sharing measure (1990)
```

Population changes for Fst calculations
By 3 climate clusters
```{r}
PopNames(msbc) <- c("Clim1", "Clim2", "Clim3") 
gpops <- c(1,1,2,2,2,2,2,2,2,2,
           2,2,2,2,2,2,2,2,2,2,
           2,2,2,2,1,1,1,1,1,1,
           1,1,3,3,3,3,3,3,3,3,
           3,3,2,2,2,2,2,2,2,2,
           2,2,2,1,1)
PopInfo(msbc) <- gpops
PopInfo(msbc)
```
PCA with climate clusters
```{r}
testmatrix <- meandistance.matrix(msbc,distmetric=Bruvo.distance)
pca1 <- cmdscale(testmatrix)
mycol <- c("#999999","#E69F00","#56B4E9","#009E73")
ggplot(pca1, aes(x=pca1[,1], y=pca1[,2], color= as.character(PopInfo(msbc)))) +
  geom_point(size=2) +
  
  stat_ellipse(geom = "polygon",
               aes(fill = as.character(PopInfo(msbc))), 
               alpha = 0.2) +
  labs(x="Principal Component 1", y="Principal Component 2") +
  labs(title = "PCA with Bruvo Distance using only B. californica") +
  scale_fill_discrete(name ="Climate", labels = c("Clim1","Clim2","Clim3")) +
  guides(color = "none") + 
  theme_bw(base_size=22)
```
By 2 Structure groups
```{r}
PopNames(msbc) <- c("Structure1","Structure2") 
gpops <- c(2,2,1,1,1,1,2,1,2,1,
           2,2,2,2,2,1,2,2,2,2,
           2,2,2,2,1,1,1,1,1,1,
           1,1,1,1,1,1,1,1,1,1,
           1,1,2,2,2,2,2,1,2,2,
           2,2,2,1,1)
PopInfo(msbc) <- gpops
```
PCA with Structure Groups
```{r}
testmatrix <- meandistance.matrix(msbc,distmetric=Bruvo.distance)
pca1 <- cmdscale(testmatrix)
mycol <- c("#999999","#E69F00","#56B4E9","#009E73")
ggplot(pca1, aes(x=pca1[,1], y=pca1[,2], color= as.character(PopInfo(msbc)))) +
  geom_point(size=2) +
  
  stat_ellipse(geom = "polygon",
               aes(fill = as.character(PopInfo(msbc))), 
               alpha = 0.2) +
  labs(x="Principal Component 1", y="Principal Component 2") +
  labs(title = "PCA with Bruvo Distance using only B. californica") +
  scale_fill_discrete(name ="Genetic Clusters", labels = c("Structure Group 1", "Structure Group 2")) +
  guides(color = "none") + 
  theme_bw(base_size=22)
```

Basic Fst calculations
```{r}
genfreqsimple <- simpleFreq(msbc)
genfreqsimple

Fst <- calcPopDiff(genfreqsimple, metric = "Fst")
Fst %>%
  round(digits=5)

#Wright's Gst (1978). States values ranging from 0-0.05 indicate “little” genetic differentiation; 
#0.05-0.15 is “moderate”, 0.15-0.25 is “great”
#microsatellites with high heterozygosity, maximum GST is often 0.1-0.2
Gst <- calcPopDiff(genfreqsimple, metric = "Gst")
Gst %>%
  round(digits=5)

#Rst is a pairwise population genetic distance that is analogous to Fst, 
#but that takes into account differences in the number of repeats between microsatellite alleles (allele size).
Rst <- calcPopDiff(genfreqsimple, metric = "Rst", object = msbc)
Rst %>%
  round(digits=5)
```

Loding in the genclone object
```{r}
cal <- read.genalex("msat_genalex_californica.csv", ploidy=4, geo=FALSE)
cal
```

Using poppr to find stats
```{r}
cal_diversity <- poppr(cal)
cal_diversity
M.tab <- mlg.table(cal)
nLoc(cal) #number of loci
ical <- informloci(cal)
ical #no uninformative loci found
nLoc(ical) #we keep all 15
```

Creating a tree
```{r}
calclust <- find.clusters(cal) #We want the lowest (most negative) BIC
calclust
calreps <- c(25, 17, 20, 21, 19, 20, 20, 20, 20, 19, 20, 19, 20, 18, 20)
calTree <- bruvo.boot(cal,replen = calreps, 
                       cutoff = 50, quiet = T)
#How do groups we found with the clustering algorithm match
#library(RColorBrewer)
cols <- brewer.pal(3, 'Set2')
plot.phylo(calTree, cex = 0.7,
           font = 2, adj = 1,
           tip.color = cols[calclust$grp],
           label.offset = 0.0002)
nodelabels(calTree$node.labels, adj = c(1.3,-0.5),
           frame = 'n',
           cex=0.8, font = 3,
           xpd = T)
axisPhylo(3)
```

Running an amova
```{r}
rho  <- poppr.amova(cal, ~Pop, within = FALSE, cutoff = .1, quiet = TRUE)
rho
```


```{r}
popNames(cal)
strata(cal)
cal2 <- as.genind(cal)
#cal2 <- gl.define.pop(cal,ind.list=c('1.03','2.04'),new='SD', verbose=NULL)
indNames(cal)
popNames(cal)
```


```{r}
cal2 <- read.genalex("msat_genalex_californica2.csv", ploidy=4, geo=FALSE)
popNames(cal2)
strata(cal2)
rho2 <-poppr.amova(cal2, ~Pop, within = FALSE, cutoff = .1, quiet = TRUE)
rho2$statphi
loc.names(cal2)
diff_test(cal, sim = TRUE, nreps = 2000)
```

```{r}
p <- c('1','1','1','1','1','1','1','1','1','1',
  '2','2','2','2','2','2','2','2','2','2',
  '3','3','3','3','3','3','3','3','3','3',
  '1','1','1','1','1','1','1','1','1','1',
  '2','2','2','2','2','2','2','2','2','2',
  '1','1','1','1','1')
pop(cal) <- p
```


```{r}
library(EcoGenetics)
library(methods)
library(ggplot2)
require(tibble)
require(poppr)
require(ade4)
require(pwr)
require(effsize)
require(sp)
require(ggmap)
require(car) 
```

Using ECOGEN Package:
```{r}
XY <- read.csv('b_coords.csv')
XY.ca <- XY %>%
  filter(Species=='californica')
XY.ca$Individual <- as.factor(XY.ca$Individual)
E <- read.csv('bca_clim.csv')
E$Individual <- as.factor(E$Individual)
E <- left_join(XY.ca, E, by="Individual")
S <- read.csv('climate.clust.csv')
S$Individual <- as.factor(S$Individual)
S <- left_join(XY.ca, S, by="Individual")
```

```{r}
ms <- read.csv("msat_genodive_conv.csv")
ms$Individual <- as.factor(ms$Individual)
ms[,c(3:62)] <- lapply(ms[,c(3:62)], sprintf, fmt="%03d")
ms$I3. <- paste(ms$I3,ms$I3.1,ms$I3.2,ms$I3.3, sep="")
ms$A1. <- paste(ms$A1,ms$A1.1,ms$A1.2,ms$A1.3, sep="")
ms$B20. <- paste(ms$B20,ms$B20.1,ms$B20.2,ms$B20.3, sep="")
ms$B11. <- paste(ms$B11,ms$B11.1,ms$B11.2,ms$B11.3, sep="")
ms$C8. <- paste(ms$C8,ms$C8.1,ms$C8.2,ms$C8.3, sep="")
ms$I14. <- paste(ms$I14,ms$I14.1,ms$I14.2,ms$I14.3, sep="")
ms$B9. <- paste(ms$B9,ms$B9.1,ms$B9.2,ms$B9.3, sep="")
ms$E9. <- paste(ms$E9,ms$E9.1,ms$E9.2,ms$E9.3, sep="")
ms$B18. <- paste(ms$B18,ms$B18.1,ms$B18.2,ms$B18.3, sep="")
ms$BF3. <- paste(ms$BF3,ms$BF3.1,ms$BF3.2,ms$BF3.3, sep="")
ms$B6. <- paste(ms$B6,ms$B6.1,ms$B6.2,ms$B6.3, sep="")
ms$BF19. <- paste(ms$BF19,ms$BF19.1,ms$BF19.2,ms$BF19.3, sep="")
ms$BF15. <- paste(ms$BF15,ms$BF15.1,ms$BF15.2,ms$BF15.3, sep="")
ms$A3. <- paste(ms$A3,ms$A3.1,ms$A3.2,ms$A3.3, sep="")
ms$B266. <- paste(ms$B266,ms$B266.1,ms$B266.2,ms$B266.3, sep="")
ms <- ms[,-c(3:62)]
names(ms) <- gsub("\\.", "", names(ms))
ms <- left_join(XY.ca, ms, by="Individual")
```

```{r}
E.ca <- E[-c(1:5)]
XY.ca<- XY.ca[-c(1:2)]
S <- S[,-c(1:4)]
#G <- G[-c(1:2),]
#G <- G[,-c(1:2)]
ms <- ms[,-c(1:5)]
```

```{r}
EG <- ecogen(XY= XY.ca, P=NULL,G=ms, E=E.ca, S=S, ploidy=4, type=c("codominant"), sep="", ncod=3, missing=0)
EG
```

```{r}
dd.genind <- EcoGenetics::ecogen2genind(EG)
ecoslot.S(EG) #Find names of cluster structures
amova.result <- poppr::poppr.amova(dd.genind, hier = ~ Region/PopID , 
                                   within=FALSE, method = "ade4")
amova.result
amova.test <- ade4::randtest(amova.result, nrepet = 999)
amova.test
```

```{r}
p <- c('1','1','2','2','2','5','5','5','6','6',
  '6','6','6','9','9','9','10','10','10','10',
  '11','11','11','11','12','12','13','13','13','13',
  '14','14','15','15','15','15','16','17','17','17',
  '18','18','20','20','20','20','21','21','22','22',
  '22','22','22','23','23')
pop(dd.genind) <- p
pop(dd.genind)
```

```{r}
summary(dd.genind)
round(pegas::hw.test(dd.genind, B=1000), digits=6)
HWE.test <- data.frame(sapply(seppop(dd.genind),
                              function(ls) pegas::hw.test(ls, B=0)[,3]))
HWE.test.chisq <- t(data.matrix(HWE.test))
HWE.test <- data.frame(sapply(seppop(dd.genind),
                              function(ls) pegas::hw.test(ls, B=1000)[,4]))
HWE.test.MC <- t(data.matrix(HWE.test))
{cat("MC permuation test (p-values):", "\n")
round(HWE.test.MC,3)}
alpha=0.05
prop.loci.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 2, mean),
                                   MC=apply(HWE.test.MC<alpha, 2, mean))
prop.loci.out.of.HWE
prop.pops.out.of.HWE <- data.frame(Chisq=apply(HWE.test.chisq<alpha, 1, mean), 
           MC=apply(HWE.test.MC<alpha, 1, mean))
prop.pops.out.of.HWE  
```

































