---
title: "Stress.Graphs"
author: "Zachary Snider"
date: "2022-11-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

#R markdown
#Saved to github repository BoecheraStress

```{r setup, include=FALSE}
#Loading in the packages
library(ggplot2)
library(tidyr)
library(hrbrthemes)
library(plotly)
library(tibble)
library(forcats)
library(scales)
library(nlme)
library(lme4)
```

Loading the Stress and Recovery Ion Leakage Data
```{r}
stress <- read.csv("jrpam_all.csv")
recovery <- read.csv("jrpam_5day.csv")
colnames(stress) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL',
                           'rel.ms','Date.Stress','Time.Stress','Fo.prime','No.',
                      'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
colnames(recovery) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL',
                      'rel.ms','Date.Stress','Time.Stress','Fo.prime','No.',
                      'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
recovery$Individual <- as.factor(recovery$Individual)
stress$Individual <- as.factor(stress$Individual)
stress$YII <- as.numeric(stress$YII)
recovery$YII <- as.numeric(recovery$YII)
summary(stress$YII)
stress <- stress %>% 
  filter(Heat.Temp == "44")
recovery <- recovery %>%
  filter(Heat.Temp == "44")
head(stress)
head(recovery)
```

Run this for californica only
```{r}
stress <- stress %>%
  filter(Species == "californica")
recovery <- recovery %>%
  filter(Species == "californica")
```


YII plots by Individual
```{r}
###Stress plot by Individual
stressplot1 <- ggplot(data=stress, aes(x=as.factor(Individual), y=YII, fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw() +
  ylab("YII") +
  xlab("Individual ID") +
  labs(title = "Initial Stress Measurements") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
stressplot1
###Recovery by Individual
recoveryplot1 <- ggplot(data=recovery, aes(x=as.factor(Individual), y=as.numeric(YII), fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw() +
  ylab("YII") +
  xlab("Individual") +
  labs(title = "5 Day Recovery Measurements") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
recoveryplot1
```

YII plots by Population
```{r}
#Stress averaged for each subpop
stressplot2 <-  ggplot(data=stress, aes(x=as.factor(PopID), y=YII, fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw(base_size=20) +
  ylab("YII") +
  xlab("Population") +
  labs(title = "Initial Stress Measurements") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
stressplot2
#Recovery averaged for each subpop
recoveryplot2 <- ggplot(data=recovery, aes(x=as.factor(PopID), y=as.numeric(YII), fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw(base_size=20) +
  ylab("YII") +
  xlab("Population") +
  labs(title = "5 Day Recovery Measurements") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
recoveryplot2
```


###PHENOTYPE DATA###

Load the data
```{r}
pheno <- read.csv("PhenotypeScores.csv")
```

For californica only
```{r}
pheno <-  pheno %>%
  filter(Species=="californica")
```


```{r}
pheno1 <- pivot_longer(pheno, cols = 2:4, names_to = "Time", values_to = "Score")
##Phenotype by Individual
pheno2 <- aggregate(Score ~ Individual + Time, 
                 data=pheno1, 
                 function(x) { 
                   c(avg=mean(x)) 
                 })

pheno2$Individual <- as.factor(pheno2$Individual)
pheno2$Time <- factor(pheno2$Time, levels=c('Pre.Stress', 'Post.Stress', 'Post.Recovery'))
```




Phenotype graph by individual
```{r}
phenoplot1 <- ggplot(data=pheno2, aes(x=Time, y=Individual,fill=Score)) +
  geom_tile(color="black", lwd=0.08) +
  scale_fill_gradient2(low="#318e1e", mid = "#ffff55", high="#ffffe9", 
                       midpoint = 1.5, na.value = "grey50") +
  theme_ipsum() +
  ylab("Individual") +
  xlab("Time") +
  labs(title = "Phenotype Score") +
  scale_x_discrete(labels=c("Pre.Stress" = "Pre Stress", "Post.Stress" = "After Stress",
                            "Post.Recovery" = "After 5 Day Recovery"))
phenoplot1
ggplotly(phenoplot1)
```

Phenotype graph by subpopulation 
```{r}
## Phenotype Scores by Subpopulation
head(pheno1)
pheno3 <- aggregate(Score ~ PopID + Time, 
                    data=pheno1, 
                    function(x) { 
                      c(avg=mean(x)) 
                    })
head(pheno3)
pheno3$PopID <- as.factor(pheno3$PopID)
pheno3$Time <- factor(pheno3$Time, levels=c('Pre.Stress', 'Post.Stress', 'Post.Recovery'))
phenoplot2 <- ggplot(data=pheno3, aes(x=Time, y=PopID,fill=Score)) +
  geom_tile(color="black", lwd=0.08) +
  scale_fill_gradient2(low="#318e1e", mid = "#ffff55", high="#ffffe9", 
                       midpoint = 1.5, na.value = "gray") +
  theme_ipsum() +
  ylab("Subpopulation") +
  xlab("Time") +
  labs(title = "Phenotype Score") +
  scale_x_discrete(labels=c("Pre.Stress" = "Pre Stress", "Post.Stress" = "After Stress",
                            "Post.Recovery" = "After 5 Day Recovery"))
phenoplot2
ggplotly(phenoplot2)
```


###ION LEAKAGE PLOT ####
loading the data
```{r}
ion <- read.csv("IonLeakage.csv")
ion$Percent.Leaked <- as.numeric(ion$Percent.Leaked)
ion$Individual <- as.factor(ion$Individual)
ion$PopID <- as.character(ion$PopID)
level_order <- c("1","2","5","6","9","10","11","12","13","14","15","17","20","21","22")
ion$PopID <- as.factor(ion$PopID)
```

For californica only
```{r}
ion <- ion %>%
  filter(Species == "californica")
```

Ion Leakage graph
```{r}
ionplot <- ggplot(ion, aes(x=factor(PopID, level = level_order), y=Percent.Leaked, fill = Region)) +
  geom_boxplot() +
  theme_bw(base_size=22) +
  xlab("Subpopulation") +
  ylab("Percent of Ions Leaked After 4 Hours")
ionplot
```

Repeating the YII graphs with only True Leaves
```{r}
#Without cotyledons, Only True Leaves
stressTL <- stress %>% 
  filter(Cotyledon.vs.TL == "True Leaf")
recoveryTL <- recovery %>%
  filter(Cotyledon.vs.TL == "True Leaf")
```

TL only Initial YII plot by subpop
```{r}
stressplotTL <- ggplot(data=stressTL, aes(x=as.factor(PopID), y=YII, fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw(base_size=20) +
  ylab("YII") +
  xlab("Individual ID") +
  labs(title = "Initial Stress Measurements (True Leaves)") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
stressplotTL
```

TL only Recovery YII plot by subpop
```{r}
recoveryplotTL <- ggplot(data=recoveryTL, aes(x=as.factor(PopID), y=as.numeric(YII), fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw(base_size=20) +
  ylab("YII") +
  xlab("Individual") +
  labs(title = "5 Day Recovery Measurements (True Leaves)") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
recoveryplotTL
```

Cotyledon only
```{r}
#Only Cotyledon
stressCTY <- stress %>% 
  filter(Cotyledon.vs.TL == "Cotyledon")
recoveryCTY <- recovery %>%
  filter(Cotyledon.vs.TL == "Cotyledon")
```

Cotyledon only initial YII plot
```{r}
stressplotCTY <- ggplot(data=stressCTY, aes(x=as.factor(PopID), y=YII, fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw() +
  ylab("YII") +
  xlab("Individual ID") +
  labs(title = "Initial Stress Measurements (Cotyledons)") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
stressplotCTY
```

Cotyledon only recovery YII plot
```{r}
recoveryplotCTY <- ggplot(data=recoveryCTY, aes(x=as.factor(PopID), y=as.numeric(YII), fill=as.factor(Region))) +
  geom_boxplot() +
  theme_bw() +
  ylab("YII") +
  xlab("Individual") +
  labs(title = "5 Day Recovery Measurements (Cotyledons)") +
  labs(fill="Heat Stress (C°)") +
  scale_fill_brewer(palette="Set2")
recoveryplotCTY
```

Comparing True Leaf and Cotyledon Initial YII values
```{r}
#True Leaf vs Cotyledon comparison
stressplot3 <- stress %>%
  drop_na(Cotyledon.vs.TL) %>%
  ggplot(aes(x=as.factor(Region), y=YII, fill=as.factor(Cotyledon.vs.TL))) +
  geom_boxplot() +
  theme_bw() +
  ylab("YII") +
  xlab("Individual ID") +
  labs(title = "Initial Stress Measurements (Cotyledons vs True Leaves)") +
  labs(fill="Type of Leaf") +
  scale_fill_manual(values = c("#8DA0CB","#66C2A5","#FC8D62"))
stressplot3
```

Comparing True Leaf and Cotyledon Recovery YII values
```{r}
recoveryplot3 <- ggplot(data=recovery, aes(x=as.factor(Region), y=as.numeric(YII),
                                           fill=as.factor(Cotyledon.vs.TL))) +
  geom_boxplot() +
  theme_bw(base_size=20) +
  ylab("YII") +
  xlab("Individual") +
  labs(title = "5 Day Recovery Measurements (Cotyledons vs True Leaves)") +
  labs(fill="Type of Leaf") +
  scale_fill_brewer(palette="Set2")
recoveryplot3
```
