---
title: "Stress2"
author: "Zachary Snider"
date: "2022-10-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

## R Markdown

Loading in the required packages
```{r setup, include=FALSE}
library(vegan)
library(FactoMineR)
library(ggplot2)
library(ggfortify)
library(tidyverse)
library(factoextra)
library(gridExtra)
library(car)
library(leaps)
library(picante)
library(hrbrthemes)
library(plotly)
library(tibble)
library(forcats)
library(scales)
library(nlme)
library(lme4)
```

Loading in the data. 
```{r}
setwd("~/Desktop/StressExperiments")
stress <- read.csv("jrpam_all.csv")
recovery <- read.csv("jrpam_5day.csv")
climate <- read.csv("boechera_climate_data.csv")
climate$PopID <- as.factor(climate$PopID)
climate$Individual <- as.factor(climate$Individual)
stress$PopID <- as.factor(stress$PopID)
stress$Individual <- as.factor(stress$Individual)
recovery$PopID <- as.factor(recovery$PopID)
recovery$Individual <- as.factor(recovery$Individual)
colnames(stress) <- c('Individual','PopID','Region','Species',
                      'Heat.Temp','Cotyledon.vs.TL',
                           'rel.ms','Date.Stress','Time.Stress','Fo.prime','No.',
                      'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
colnames(recovery) <- c('Individual','PopID','Region','Species','Heat.Temp',
                        'Cotyledon.vs.TL', 'rel.ms','Date.Stress','Time.Stress',
                        'Fo.prime','No.', 'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
stress.clim <- left_join(stress, climate, by=c("Individual","Region","PopID","Species"))
colnames(stress.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') #Renaming columns for legibility
recovery.clim <- left_join(recovery, climate, by=c("Individual","Region","PopID","Species"))
colnames(recovery.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') 
```

Subsetting Data for only CALIFORNICA
```{r}
stress.ca <- stress %>%
  filter(Species == "californica")
recovery.ca <- recovery %>%
  filter(Species == "californica")
stress.clim.ca <- stress.clim %>%
  filter(Species == "californica")
recovery.clim.ca <- recovery.clim %>%
  filter(Species == "californica")

```

### 5 DAY RECOVERY CHLOROPHYLL FLUORESCENCE DATA ###

Nested ANOVA to test the variance of YII. 
The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#First create a new column with the variances of each individual
recovery.ca.2 <- aggregate(YII ~ Individual, 
                    data=recovery, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(recovery.ca.2) <- c('Individual','YII.var')
recovery.ca.3 <- left_join(recovery.ca.2, recovery, by=c("Individual"))
recovery.ca.3 <- recovery.ca.3[!duplicated(recovery.ca.3$Individual), ]
head(recovery.ca.3)
#Make the nested anova
nest_r <- aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region / factor(recovery.ca.3$PopID) / factor(recovery.ca.3$Individual))
summary(nest_r)
#TukeyHSD(nest)

nest2_r <- aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region + recovery.ca.3$Region / factor(recovery.ca.3$PopID) + (recovery.ca.3$Region / factor(recovery.ca.3$PopID) / factor(recovery.ca.3$Individual)))
summary(nest2_r)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_r <- lmer(YII ~ Region + (1|Individual), data = recovery.ca)
summary(mixed.lmer_r)
plot(mixed.lmer_r)
qqnorm(resid(mixed.lmer_r)); qqline(resid(mixed.lmer_r))
#Variance for Individual is 0.0216. Divide this by total variance. 
0.0216 / (0.0216 + 0.0377) # =0.36
#Individual explains 36% of the variance left over after the variance explained by our fixed effects.
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
rmod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca, nvmax=10)
sum.rmod <- summary(rmod.fit)
sum.rmod
plot(sum.rmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.rmod$adjr2) #8
points(adj_r2_max, sum.rmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
rmod <- lm(YII ~ tmean + Elevation + tmax  + tdmean + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca)
summary(rmod) #adjusted r squared = 0.341
Anova(rmod, type=2)
plot(rmod)
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=recovery.clim.ca, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=recovery.clim.ca, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=recovery.clim.ca, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=recovery.clim.ca, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=recovery.clim.ca, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=recovery.clim.ca, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=recovery.clim.ca, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```

### INITIAL CHLOROPHYLL FLUORESCENCE DATA ###

Nested ANOVA to test the variance of YII. 
The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#First create a new column with the variances of each individual
stress.ca.2 <- aggregate(YII ~ Individual, 
                    data=stress, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(stress.ca.2) <- c('Individual','YII.var')
stress.ca.3 <- left_join(stress.ca.2, stress, by=c("Individual"))
stresss.ca.3 <- stress.ca.3[!duplicated(stress.ca.3$Individual), ]
head(stress.ca.3)
#Make the nested anova
nest_s <- aov(stress.ca.3$YII.var ~ stress.ca.3$Region / factor(stress.ca.3$PopID) / factor(stress.ca.3$Individual))
summary(nest_s)
#TukeyHSD(nest)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_s <- lmer(YII ~ Region + (1|Individual), data = stress.ca)
summary(mixed.lmer_s)
plot(mixed.lmer_s)
qqnorm(resid(mixed.lmer_s)); qqline(resid(mixed.lmer_s))
#Variance for Individual is 0.0216. Divide this by total variance. 
0.0122 / (0.0122 + 0.0272) # =0.31
#Individual explains 31% of the variance left over after the variance explained by our fixed effects.
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
rmod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca, nvmax=10)
sum.rmod <- summary(rmod.fit)
sum.rmod
plot(sum.rmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.rmod$adjr2) #8
points(adj_r2_max, sum.rmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
rmod <- lm(YII ~ tmean + Elevation + tmax  + tdmean + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca)
summary(rmod) #adjusted r squared = 0.341
Anova(rmod, type=2)
plot(rmod)
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=recovery.clim.ca, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=recovery.clim.ca, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=recovery.clim.ca, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=recovery.clim.ca, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=recovery.clim.ca, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=recovery.clim.ca, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=recovery.clim.ca, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```




```{r}


```


```{r}


```

### ION LEAKAGE ###
```{r}
ion <- read.csv("IonLeakage.csv")
ion$Percent.Leaked <- as.numeric(ion$Percent.Leaked)
ion$Individual <- as.factor(ion$Individual)
ion$PopID <- as.factor(ion$PopID)
ion.ca <- ion %>%
  filter(Species == "californica")

ioncaplot <- ggplot(ion.ca, aes(x=PopID, y=Percent.Leaked, fill = Region)) +
  geom_boxplot() +
  theme_bw() +
  xlab("Subpopulation") +
  ylab("Percent of Ions Leaked After 4 Hours")
ioncaplot

```

### PHENOTYPE DATA ###
```{r}
pheno <- read.csv("PhenotypeScores.csv")
pheno$Individual <- as.factor(pheno$Individual)
pheno.ca <- pheno %>%
  filter(Species == "californica")
pheno.ca1 <- pivot_longer(pheno.ca, cols = 2:4, names_to = "Time", values_to = "Score")
pheno.ca2 <- aggregate(Score ~ PopID + Time, 
                    data=pheno.ca1, 
                    function(x) { 
                      c(avg=mean(x)) 
                    })
pheno.ca2$PopID <- as.factor(pheno.ca2$PopID)
pheno.ca2$Time <- factor(pheno.ca2$Time, levels=c('Pre.Stress', 'Post.Stress', 'Post.Recovery'))
phenoplot2 <- ggplot(data=pheno.ca2, aes(x=Time, y=PopID,fill=Score)) +
  geom_tile(color="black", lwd=0.08) +
  scale_fill_gradient2(low="#318e1e", mid = "#ffff55", high="#ffffe9", 
                       midpoint = 1.5, na.value = "gray") +
  theme_ipsum() +
  ylab("Subpopulation") +
  xlab("Time") +
  labs(title = "Phenotype Score") +
  scale_x_discrete(labels=c("Pre.Stress" = "Pre Stress", "Post.Stress" = "After Stress",
                            "Post.Recovery" = "After 5 Day Recovery"))
phenoplot2
ggplotly(phenoplot2)
```


```{r}


```


```{r}


```



























