---
title: "Stress.YII.TL"
author: "Zachary Snider"
date: "2022-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

Loading in the required packages
```{r setup, include=FALSE}
library(vegan)
library(FactoMineR)
library(ggplot2)
library(ggfortify)
library(tidyverse)
library(factoextra)
library(gridExtra)
library(car)
library(leaps)
library(picante)
library(hrbrthemes)
library(plotly)
library(tibble)
library(forcats)
library(scales)
library(nlme)
library(lme4)
```

Loading in the data. 
```{r}
setwd("~/Desktop/StressExperiments")
stress <- read.csv("jrpam_all.csv")
recovery <- read.csv("jrpam_5day.csv")
climate <- read.csv("boechera_climate_data.csv")
climate$PopID <- as.factor(climate$PopID)
climate$Individual <- as.factor(climate$Individual)
stress$PopID <- as.factor(stress$PopID)
stress$Individual <- as.factor(stress$Individual)
recovery$PopID <- as.factor(recovery$PopID)
recovery$Individual <- as.factor(recovery$Individual)
colnames(stress) <- c('Individual','PopID','Region','Species',
                      'Heat.Temp','Cotyledon.vs.TL',
                           'rel.ms','Date.Stress','Time.Stress','Fo.prime','No.',
                      'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
colnames(recovery) <- c('Individual','PopID','Region','Species','Heat.Temp',
                        'Cotyledon.vs.TL', 'rel.ms','Date.Stress','Time.Stress',
                        'Fo.prime','No.', 'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
stress.clim <- left_join(stress, climate, by=c("Individual","Region","PopID","Species"))
colnames(stress.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') #Renaming columns for legibility
recovery.clim <- left_join(recovery, climate, by=c("Individual","Region","PopID","Species"))
colnames(recovery.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') 

```

Subsetting Data for only TRUE LEAVES
```{r}
stress.clim.ca <- stress.clim %>%
  filter(Species == "californica")
recovery.clim.ca <- recovery.clim %>%
  filter(Species == "californica")
stress.clim.ca.tl <- stress.clim.ca %>%
  filter(Cotyledon.vs.TL == "True Leaf")
recovery.clim.ca.tl <- recovery.clim.ca %>%
  filter(Cotyledon.vs.TL == "True Leaf")
```

Creating a variance and a mean YII group
```{r}
recovery.clim.ca.tl.var <- aggregate(YII ~ Individual, 
                    data=recovery.clim.ca.tl, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(recovery.clim.ca.tl.var) <- c('Individual','YII.var')
recovery.clim.ca.tl.2 <- left_join(recovery.clim.ca.tl.var, recovery.clim.ca.tl, by=c("Individual"))
recovery.clim.ca.tl.3 <- recovery.clim.ca.tl.2[!duplicated(recovery.clim.ca.tl.2$Individual), ]
head(recovery.clim.ca.tl.2)
recovery.clim.ca.tl.2 <- recovery.clim.ca.tl.2[,-c(8,9,10,11,12,23)]
recovery.clim.ca.tl.mean <- aggregate(YII ~ Individual, 
                    data=recovery.clim.ca.tl, 
                    function(x) { 
                      c(avg=mean(x)) 
                    })
colnames(recovery.clim.ca.tl.mean) <- c('Individual','YII.mean')
recovery.clim.ca.tl.2 <- left_join(recovery.clim.ca.tl.mean, recovery.clim.ca.tl.2, by=c("Individual"))
recovery.clim.ca.tl.3 <- recovery.clim.ca.tl.2[!duplicated(recovery.clim.ca.tl.2$Individual), ]
head(recovery.clim.ca.tl.2)
recovery.clim.ca.tl.3 <- recovery.clim.ca.tl.3[,-c(8,9,10,11,12,23)]

```


### 5 DAY RECOVERY CHLOROPHYLL FLUORESCENCE DATA ###

First testing for interactions
```{r}
m0 <- lm(YII.var ~ Region, data=recovery.clim.ca.tl.2)
m1 <- lm(YII.var ~ Region+PopID, data=recovery.clim.ca.tl.2)
anova(m1,m0) # test for difference in mean(PopID) after control for Rregion
m2 <- lm(YII.var ~ Region*PopID, data=recovery.clim.ca.tl.2) 
anova(m2, m1) # test for interaction: different slopes in the two Conditions

```

Nested ANOVA to test the variance of YII. 
The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#Make the nested anova
nest <- aov(recovery.clim.ca.tl.2$YII.var ~ recovery.clim.ca.tl.2$Region / factor(recovery.clim.ca.tl.2$PopID) / factor(recovery.clim.ca.tl.2$Individual))
summary(nest)
nest
#TukeyHSD(nest)

```

Linear Mixed Effects Model
```{r}
mixed.lmer_r <- lmer(YII.var ~ Region + (1|PopID:Region), data = recovery.clim.ca.tl.3)
#mixed.lmer_r2 <- lmer(YII.var ~ Region + (1|PopID) + (Region|Individual), data = recovery.clim.ca.2)
#mixed.lme_r <- lme(YII.var ~ Region, random = (1|PopID), data=recovery.ca.3)
#anova(mixed.lmer_r, mixed.lmer_r2)
summary(mixed.lmer_r)
#summary(mixed.lmer_r2)
plot(mixed.lmer_r)
qqnorm(resid(mixed.lmer_r)); qqline(resid(mixed.lmer_r))
0.0011 / (0.0011+0.000405) #=0.73
#PopID explains 73% of the variance left over
#Variance for Individual is 0.0216. Divide this by total variance. 
0.001432 / (0.001432 + 0.000244) 

library(effects)
plot(allEffects(mixed.lmer_r))
anova(mixed.lmer_r)
#library(broom)
#tidy(mixed.lmer_r, effects = "fixed")
#glance(mixed.lmer_r)
confint(mixed.lmer_r)
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
rmod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca.tl.2, nvmax=10)
sum.rmod <- summary(rmod.fit)
sum.rmod
plot(sum.rmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.rmod$adjr2) #9
points(adj_r2_max, sum.rmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca.tl.2, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca.tl.2, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
rmod <- lm(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + solclear, data=recovery.clim.ca.tl.2)
summary(rmod) #adjusted r squared = 0.323
Anova(rmod, type=2)
par(mfrow=c(2,2))
plot(rmod)
par(mfrow=c(1,1))
```

Checking correlation of variables. 
```{r}
head(recovery.clim.ca.tl.2)
cor.table(recovery.clim.ca.tl.2[,c(18:25,27,28)])
cor_data <- cor((recovery.clim.ca.tl.2[,c(18:25,27,28)]))
cor_data
cor_matrix_rm <- cor_data                # Modify correlation matrix
cor_matrix_rm[upper.tri(cor_matrix_rm)] <- 0
diag(cor_matrix_rm) <- 0
cor_matrix_rm
#To remove highly correlated variables
recovery.clim.ca.tl.4 <- recovery.clim.ca[ , !apply(cor_matrix_rm,2,function(x) any(x > 0.99))]
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```