---
title: "Stress2"
author: "Zachary Snider"
date: "2022-10-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

## R Markdown

Stress and Climate data for Southern California Boechera populations collected by Zach Snider. Variables include: Invididual, subpopulation (PopID), regional population (Region), Temperature of heat stress experiment (Heat.Temp), Cotyledon or True Leaf measurements (Cotyledon.vs.TL), Relative time, Date of heat stress, time of heat stress, type of stress measurement, count of measurement for that session, 1:F, 1:Fm, 1:Temp, YII,, 1:Fo,, 1:Fm, 1:Fv/Fm , Latitude, Longitude, Elevation, Date.Found, mean temperature, average max temp, average min temp, average precipitation, average max vapor pressure deficit, average min vapor pressure deficit, aspect, average solar transmittance on a clear day, and average solar transmittance on a cloudy day. 

Elevation along with the 9 climate variables are potential explanatory variables, and could be appropriate for a multiple regression approach to predicting YII (quantum yield), but 10
predictors is a lot. So it may be preferable to reduce the number of explanatory
variables using PCA to come up with a few important explanatory components.

Loading in the required packages
```{r setup, include=FALSE}
library(vegan)
library(FactoMineR)
library(ggplot2)
library(ggfortify)
library(tidyverse)
library(factoextra)
library(gridExtra)
library(car)
library(leaps)
library(picante)
library(hrbrthemes)
library(plotly)
library(tibble)
library(forcats)
library(scales)
library(nlme)
library(lme4)
```

Loading in the data. 
```{r}
#setwd("~/Desktop/StressExperiments")
#setwd("C:/Users/Daniel/Desktop/BoecheraStress-main")
stress <- read.csv("jrpam_all.csv")
recovery <- read.csv("jrpam_5day.csv")
climate <- read.csv("boechera_climate_data.csv")
climate$PopID <- as.factor(climate$PopID)
climate$Individual <- as.factor(climate$Individual)
stress$PopID <- as.factor(stress$PopID)
stress$Individual <- as.factor(stress$Individual)
recovery$PopID <- as.factor(recovery$PopID)
recovery$Individual <- as.factor(recovery$Individual)
colnames(stress) <- c('Individual','PopID','Region','Species',
                      'Heat.Temp','Cotyledon.vs.TL',
                           'rel.ms','Date.Stress','Time.Stress','Fo.prime','No.',
                      'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
colnames(recovery) <- c('Individual','PopID','Region','Species','Heat.Temp',
                        'Cotyledon.vs.TL', 'rel.ms','Date.Stress','Time.Stress',
                        'Fo.prime','No.', 'F','Fm.','Temp','YII','Fo','Fm','Fv.Fm')
stress.clim <- left_join(stress, climate, by=c("Individual","Region","PopID","Species"))
colnames(stress.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') #Renaming columns for legibility
recovery.clim <- left_join(recovery, climate, by=c("Individual","Region","PopID","Species"))
colnames(recovery.clim) <- c('Individual','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','rel.ms','Date.Stress','Time.Stress','Fo.prime','No.','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') 

```

Subsetting Data for only CALIFORNICA
```{r}
stress.ca <- stress %>%
  filter(Species == "californica")
recovery.ca <- recovery %>%
  filter(Species == "californica")
stress.clim.ca <- stress.clim %>%
  filter(Species == "californica")
recovery.clim.ca <- recovery.clim %>%
  filter(Species == "californica")

```

### 5 DAY RECOVERY CHLOROPHYLL FLUORESCENCE DATA ###

Nested ANOVA to test the variance of YII. 
```{r}
#First create a new column with the variances of each individual
recovery.ca.2 <- aggregate(YII ~ Individual, 
                    data=recovery, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(recovery.ca.2) <- c('Individual','YII.var')
recovery.ca.3 <- left_join(recovery.ca.2, recovery.ca, by=c("Individual"))
recovery.ca.3 <- recovery.ca.3[!duplicated(recovery.ca.3$Individual), ]
recovery.clim.ca.2 <- left_join(recovery.ca.3, climate, by=c("Individual","Region","PopID","Species"))
head(recovery.clim.ca.2)
recovery.clim.ca.2 <- recovery.clim.ca.2[,-c(8,9,10,11,12,23)]
colnames(recovery.clim.ca.2) <- c('Individual','YII.var','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') 
recovery.clim.ca.2 <- recovery.clim.ca.2 %>%
  filter(Species == "californica")
recovery.ca.3 <- recovery.ca.3 %>%
  filter(Species == "californica")
```

First testing for interactions
```{r}
m0 <- lm(YII.var ~ Region, data=recovery.clim.ca.2)
m1 <- lm(YII.var ~ Region+PopID, data=recovery.clim.ca.2)
anova(m1,m0) # test for difference in mean(PopID) after control for Rregion
m2 <- lm(YII.var ~ Region*PopID, data=recovery.clim.ca.2) 
anova(m2, m1) # test for interaction: different slopes in the two Conditions

```


The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#Make the nested anova
nest_r <- aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region / factor(recovery.ca.3$PopID) / factor(recovery.ca.3$Individual))
summary(nest_r)
nest_r
#TukeyHSD(nest)

nest2_r <- aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region + recovery.ca.3$Region / factor(recovery.ca.3$PopID) + (recovery.ca.3$Region / factor(recovery.ca.3$PopID) / factor(recovery.ca.3$Individual)))
summary(nest2_r)

nest3_r <-  aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region * factor(recovery.ca.3$PopID) * factor(recovery.ca.3$Individual))
summary(nest3_r)
anova(nest3_r)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_r <- lmer(YII.var ~ Region + (1|PopID:Region), data = recovery.clim.ca.2)
#mixed.lmer_r2 <- lmer(YII.var ~ Region + (1|PopID) + (Region|Individual), data = recovery.clim.ca.2)
#mixed.lme_r <- lme(YII.var ~ Region, random = (1|PopID), data=recovery.ca.3)
#anova(mixed.lmer_r, mixed.lmer_r2)
summary(mixed.lmer_r)
#summary(mixed.lmer_r2)
plot(mixed.lmer_r)
qqnorm(resid(mixed.lmer_r)); qqline(resid(mixed.lmer_r))
0.000294 / (0.000204+0.000181) #=0.76
#PopID explains 76% of the variance left over
#Variance for Individual is 0.0216. Divide this by total variance. 
0.014 / (0.014 + 0.0131 + 0.0377) # =0.22
#Individual explains 22% of the variance left over after the variance explained by our fixed effects.
0.0131 / (0.014 + 0.0131 + 0.0377) # =0.2
#Pop ID explains 20% of the variance.

library(effects)
plot(allEffects(mixed.lmer_r))
anova(mixed.lmer_r)
#library(broom)
#tidy(mixed.lmer_r, effects = "fixed")
#glance(mixed.lmer_r)
confint(mixed.lmer_r)
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
rmod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca, nvmax=10)
sum.rmod <- summary(rmod.fit)
sum.rmod
plot(sum.rmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.rmod$adjr2) #9
points(adj_r2_max, sum.rmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
rmod <- lm(YII ~ tmean + Elevation + tmax  + tdmean + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca)
summary(rmod) #adjusted r squared = 0.341
Anova(rmod, type=2)
par(mfrow=c(2,2))
plot(rmod)
par(mfrow=c(1,1))
```

Checking correlation of variables. 
```{r}
cor.table(recovery.clim.ca[,c(21,23:27,29,31,32)])
rmod <- lm(YII ~ tmean)
cor_data <- cor((recovery.clim.ca[,c(21,23:27,29,31,32)]))
cor_data
cor_matrix_rm <- cor_data                # Modify correlation matrix
cor_matrix_rm[upper.tri(cor_matrix_rm)] <- 0
diag(cor_matrix_rm) <- 0
cor_matrix_rm
#To remove highly correlated variables
recovery.clim.ca.4 <- recovery.clim.ca[ , !apply(cor_matrix_rm,2,function(x) any(x > 0.99))]
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=recovery.clim.ca, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=recovery.clim.ca, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=recovery.clim.ca, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=recovery.clim.ca, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=recovery.clim.ca, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=recovery.clim.ca, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=recovery.clim.ca, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```

### INITIAL CHLOROPHYLL FLUORESCENCE DATA ###

Nested ANOVA to test the variance of YII. 
The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#First create a new column with the variances of each individual
stress.ca.2 <- aggregate(YII ~ Individual, 
                    data=stress, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(stress.ca.2) <- c('Individual','YII.var')
stress.ca.3 <- left_join(stress.ca.2, stress, by=c("Individual"))
stresss.ca.3 <- stress.ca.3[!duplicated(stress.ca.3$Individual), ]
head(stress.ca.3)
#Make the nested anova
nest_s <- aov(stress.ca.3$YII.var ~ stress.ca.3$Region / factor(stress.ca.3$PopID) / factor(stress.ca.3$Individual))
summary(nest_s)
#TukeyHSD(nest)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_s <- lmer(YII ~ Region + (1|Individual), data = stress.ca)
summary(mixed.lmer_s)
plot(mixed.lmer_s)
qqnorm(resid(mixed.lmer_s)); qqline(resid(mixed.lmer_s))
#Variance for Individual is 0.0216. Divide this by total variance. 
0.0122 / (0.0122 + 0.0272) # =0.31
#Individual explains 31% of the variance left over after the variance explained by our fixed effects.
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
smod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=stress.clim.ca, nvmax=10)
sum.smod <- summary(smod.fit)
sum.smod
plot(sum.smod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.smod$adjr2) #6
points(adj_r2_max, sum.smod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = stress.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = stress.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
smod <- lm(YII ~Elevation + tmax + PPT + vpdmin + solclear + soltrans, data=stress.clim.ca)
summary(smod) #adjusted r squared = 0.341
Anova(smod, type=2)
plot(smod)
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=stress.clim.ca, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=stress.clim.ca, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=stress.clim.ca, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=stress.clim.ca, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=stress.clim.ca, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=stress.clim.ca, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=stress.clim.ca, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=stress.clim.ca, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=stress.clim.ca, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=stress.clim.ca, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```




```{r}


```


```{r}


```

### ION LEAKAGE ###
```{r}
ion <- read.csv("IonLeakage.csv")
#Merging the ion leakage data and climate data
ion.clim <- left_join(ion, climate, by=c("Individual","Region","PopID","Species"))
colnames(ion.clim) <- c('Individual','PopID','Region','Species','Date','0H','4H','Total.24.Freeze','Percent.Leaked','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') #Renaming columns for legibility
ion.clim$Percent.Leaked <- as.numeric(ion.clim$Percent.Leaked)
ion.clim$Individual <- as.factor(ion.clim$Individual)
ion.clim$PopID <- as.factor(ion.clim$PopID)
ion.clim.ca <- ion.clim %>%
  filter(Species == "californica")

ioncaplot <- ggplot(ion.clim.ca, aes(x=PopID, y=Percent.Leaked, fill = Region)) +
  geom_boxplot() +
  theme_bw() +
  xlab("Subpopulation") +
  ylab("Percent of Ions Leaked After 4 Hours")
ioncaplot
```

Nested ANOVA to test the variance of Ion Leakage. 
The model is Ion Leakage ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#Make the nested anova
nest_i <- aov(ion.clim.ca$Percent.Leaked ~ ion.clim.ca$Region / factor(ion.clim.ca$PopID) / factor(ion.clim.ca$Individual))
summary(nest_i)
#TukeyHSD(nest)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_i <- lmer(Percent.Leaked ~ Region + (1|Individual), data = ion.clim.ca)
summary(mixed.lmer_i)
plot(mixed.lmer_i)
qqnorm(resid(mixed.lmer_i)); qqline(resid(mixed.lmer_i))
#Variance for Individual is 0.0. Divide this by total variance. 
0.0 / (0.0 + 34.4) # =0.0
#Individual explains 31% of the variance left over after the variance explained by our fixed effects.
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
imod.fit <- regsubsets(Percent.Leaked ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=ion.clim.ca, nvmax=10)
sum.imod <- summary(imod.fit)
sum.imod
plot(sum.imod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.imod$adjr2) #6
points(adj_r2_max, sum.imod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(Percent.Leaked~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = ion.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(Percent.Leaked~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = ion.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
imod <- lm(Percent.Leaked ~ vpdmax + PPT + solclear, data=ion.clim.ca)
summary(imod) #adjusted r squared = 0.341
Anova(imod, type=2)
par(mfrow = c(2, 2))
plot(imod)
par(mfrow = c(1, 1))
```

Plotting ion leakage against climatic variables
```{r}
p1 <- ggplot(data=ion.clim.ca, aes(x=tmean, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=ion.clim.ca, aes(x=tmax, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=ion.clim.ca, aes(x=tmin, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=ion.clim.ca, aes(x=Elevation, y=Percent.Leaked)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=ion.clim.ca, aes(x=PPT, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=ion.clim.ca, aes(x=tdmean, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=ion.clim.ca, aes(x=vpdmax, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=ion.clim.ca, aes(x=vpdmin, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=ion.clim.ca, aes(x=solclear, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=ion.clim.ca, aes(x=soltrans, y=Percent.Leaked)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```

### PHENOTYPE DATA ###
```{r}
pheno <- read.csv("PhenotypeScores.csv")
pheno$Individual <- as.factor(pheno$Individual)
pheno.ca <- pheno %>%
  filter(Species == "californica")
pheno.ca.1 <- pivot_longer(pheno.ca, cols = 2:4, names_to = "Time", values_to = "Score")
pheno.ca.2 <- aggregate(Score ~ PopID + Time, 
                    data=pheno.ca.1, 
                    function(x) { 
                      c(avg=mean(x)) 
                    })
pheno.ca.2$PopID <- as.factor(pheno.ca.2$PopID)
colnames(pheno.ca.2) <- c('PopID','Time','Score.pop.mean')
pheno.ca.2$Time <- factor(pheno.ca.2$Time, levels=c('Pre.Stress', 'Post.Stress', 'Post.Recovery'))
phenoplot2 <- ggplot(data=pheno.ca.2, aes(x=Time, y=PopID,fill=Score.pop.mean)) +
  geom_tile(color="black", lwd=0.08) +
  scale_fill_gradient2(low="#318e1e", mid = "#ffff55", high="#ffffe9", 
                       midpoint = 1.5, na.value = "gray") +
  theme_ipsum() +
  ylab("Subpopulation") +
  xlab("Time") +
  labs(title = "Phenotype Score") +
  scale_x_discrete(labels=c("Pre.Stress" = "Pre Stress", "Post.Stress" = "After Stress",
                            "Post.Recovery" = "After 5 Day Recovery"))
phenoplot2
ggplotly(phenoplot2)
```


```{r}
#Create a new column with the variances of each individual
pheno.ca.3 <- aggregate(Score ~ Individual + Time, 
                    data=pheno.ca.1, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(pheno.ca.3) <- c('Individual','Time','Score.var')
pheno.ca.3 <- left_join(pheno.ca.3, pheno.ca.1, by=c('Individual','Time'))
pheno.ca.3$Individual <- as.factor(pheno.ca.3$Individual)
pheno.ca.3 <- pheno.ca.3[!duplicated(pheno.ca.3[,c('Individual','Time')]), ]
pheno.ca.4 <- aggregate(Score ~ Individual + Time, 
                    data=pheno.ca.1, 
                    function(x) { 
                      c(avg=mean(x)) 
                    })
colnames(pheno.ca.4) <- c('Individual','Time','Score.mean')
pheno.ca.4 <- left_join(pheno.ca.3, pheno.ca.4, by=c('Individual','Time'))
pheno.ca.4$Individual <- as.factor(pheno.ca.4$Individual)
pheno.ca.4$PopID <- as.factor(pheno.ca.4$PopID)
pheno.ca.4 <- pheno.ca.4[,-7]
pheno.clim.ca <- left_join(pheno.ca.4, climate, by=c("Individual","Region","PopID","Species"))
colnames(pheno.clim.ca) <- c('Individual','Time','Score.var','PopID','Region','Species','Score.mean','Latitude','Longitude','Elevation','Date.Found','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') #Renaming columns for legibility
```

Nested ANOVA to test the variance of Phenotype Score. 
The model is Score ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#Make the nested anova
nest_p <- aov(pheno.clim.ca$Score.mean ~ pheno.clim.ca$Region / factor(pheno.clim.ca$PopID) / factor(pheno.clim.ca$Individual))
summary(nest_p)
#TukeyHSD(nest)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_p <- lmer(Score.mean ~ Region + (1|Individual), data = pheno.clim.ca)
summary(mixed.lmer_p)
plot(mixed.lmer_p)
qqnorm(resid(mixed.lmer_p)); qqline(resid(mixed.lmer_p))
#Variance for Individual is 0.0. Divide this by total variance. 
0.0 / (0.0 + 0.794) # =0.0
#Individual explains 0% of the variance left over after the variance explained by our fixed effects.
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
pmod.fit <- regsubsets(Score.mean ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=pheno.clim.ca, nvmax=10)
sum.pmod <- summary(pmod.fit)
sum.pmod
plot(sum.pmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.pmod$adjr2) #7
points(adj_r2_max, sum.pmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(Score.mean ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = pheno.clim.ca, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(Score.mean ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = pheno.clim.ca, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
pmod <- lm(Score.mean ~ tmean + tdmean + vpdmax + tmin + PPT + solclear + soltrans, data=pheno.clim.ca)
summary(pmod) #adjusted r squared = 0.0886
Anova(pmod, type=2)
par(mfrow = c(2, 2))
plot(pmod)
par(mfrow = c(1, 1))
```

Plotting ion leakage against climatic variables
```{r}
p1 <- ggplot(data=pheno.clim.ca, aes(x=tmean, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=pheno.clim.ca, aes(x=tmax, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=pheno.clim.ca, aes(x=tmin, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=pheno.clim.ca, aes(x=Elevation, y=Score.mean)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=pheno.clim.ca, aes(x=PPT, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=pheno.clim.ca, aes(x=tdmean, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=pheno.clim.ca, aes(x=vpdmax, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=pheno.clim.ca, aes(x=vpdmin, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=pheno.clim.ca, aes(x=solclear, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=pheno.clim.ca, aes(x=soltrans, y=Score.mean)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```





```{r}
structure.ca <- read.csv("Structure_group_ca.csv")
structure.ca$Individual <- as.factor(structure.ca$Individual)
recovery.clim.ca.3 <- left_join(recovery.clim.ca.2, structure.ca, by=c("Individual"))
recovery.clim.ca.3$K2 <- as.factor(recovery.clim.ca.3$K2)
recovery.clim.ca.3$K3 <- as.factor(recovery.clim.ca.3$K3)
```


```{r}
p1 <- ggplot(data=recovery.clim.ca.3, aes(x=tmean, y=YII, color=K2)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca.3, aes(x=tmax, y=YII, color=K2)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca.3, aes(x=tmin, y=YII, color=K2)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca.3, aes(x=Elevation, y=YII, color=K2)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)

```

```{r}
p1 <- ggplot(data=recovery.clim.ca.3, aes(x=tmean, y=YII, color=K3)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca.3, aes(x=tmax, y=YII, color=K3)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca.3, aes(x=tmin, y=YII, color=K3)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca.3, aes(x=Elevation, y=YII, color=K3)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)

```














Subsetting Data for only TRUE LEAVES
```{r}
stress.ca <- stress %>%
  filter(Species == "californica")
recovery.ca <- recovery %>%
  filter(Species == "californica")
stress.clim.ca <- stress.clim %>%
  filter(Species == "californica")
recovery.clim.ca <- recovery.clim %>%
  filter(Species == "californica")

stress.ca.tl <- stress.ca %>%
  filter(Cotyledon.vs.TL == "True Leaf")
recovery.ca.tl <- recovery.ca %>%
  filter(Cotyledon.vs.TL == "True Leaf")
```

### 5 DAY RECOVERY CHLOROPHYLL FLUORESCENCE DATA ###

Nested ANOVA to test the variance of YII. 
```{r}
#First create a new column with the variances of each individual
recovery.ca.tl.2 <- aggregate(YII ~ Individual, 
                    data=recovery.ca.tl, 
                    function(x) { 
                      c(avg=var(x)) 
                    })
colnames(recovery.ca.tl.2) <- c('Individual','YII.var')
recovery.ca.tl.3 <- left_join(recovery.ca.tl.2, recovery.ca.tl, by=c("Individual"))
recovery.ca.tl.3 <- recovery.ca.tl.3[!duplicated(recovery.ca.tl.3$Individual), ]
recovery.clim.ca.tl.2 <- left_join(recovery.ca.tl.3, climate, by=c("Individual","Region","PopID","Species"))
head(recovery.clim.ca.tl.2)
recovery.clim.ca.tl.2 <- recovery.clim.ca.tl.2[,-c(8,9,10,11,12,23)]
colnames(recovery.clim.ca.tl.2) <- c('Individual','YII.var','PopID','Region','Species','Heat.Temp','Cotyledon.vs.TL','F','Fm.','Temp','YII','Fo','Fm','Fv.Fm','Latitude','Longitude','Elevation','tmean','tmax','tmin','PPT','tdmean','vpdmax','vpdmin','aspect','solclear','soltrans') 
recovery.clim.ca.tl.2 <- recovery.clim.ca.tl.2 %>%
  filter(Species == "californica")
recovery.ca.tl.3 <- recovery.ca.tl.3 %>%
  filter(Species == "californica")
```

First testing for interactions
```{r}
m0 <- lm(YII.var ~ Region, data=recovery.clim.ca.2)
m1 <- lm(YII.var ~ Region+PopID, data=recovery.clim.ca.2)
anova(m1,m0) # test for difference in mean(PopID) after control for Rregion
m2 <- lm(YII.var ~ Region*PopID, data=recovery.clim.ca.2) 
anova(m2, m1) # test for interaction: different slopes in the two Conditions

```


The model is YII ~ Region + Pop(Region) + Individual(Pop(Region))
Individual is a random effect. 
```{r}
#Make the nested anova
nest_rtl <- aov(recovery.ca.tl.3$YII.var ~ recovery.ca.tl.3$Region / factor(recovery.ca.tl.3$PopID) / factor(recovery.ca.tl.3$Individual))
summary(nest_rtl)
nest_rtl
#TukeyHSD(nest)

nest2_r <- aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region + recovery.ca.3$Region / factor(recovery.ca.3$PopID) + (recovery.ca.3$Region / factor(recovery.ca.3$PopID) / factor(recovery.ca.3$Individual)))
summary(nest2_r)

nest3_r <-  aov(recovery.ca.3$YII.var ~ recovery.ca.3$Region * factor(recovery.ca.3$PopID) * factor(recovery.ca.3$Individual))
summary(nest3_r)
anova(nest3_r)
```

Linear Mixed Effects Model
```{r}
mixed.lmer_r <- lmer(YII.var ~ Region + (1|PopID:Region), data = recovery.clim.ca.tl.2)
#mixed.lmer_r2 <- lmer(YII.var ~ Region + (1|PopID) + (Region|Individual), data = recovery.clim.ca.2)
#mixed.lme_r <- lme(YII.var ~ Region, random = (1|PopID), data=recovery.ca.3)
#anova(mixed.lmer_r, mixed.lmer_r2)
summary(mixed.lmer_r)
#summary(mixed.lmer_r2)
plot(mixed.lmer_r)
qqnorm(resid(mixed.lmer_r)); qqline(resid(mixed.lmer_r))
0.0011 / (0.0011+0.000405) #=0.73
#PopID explains 73% of the variance left over
#Variance for Individual is 0.0216. Divide this by total variance. 
0.014 / (0.014 + 0.0131 + 0.0377) # =0.22
#Individual explains 22% of the variance left over after the variance explained by our fixed effects.
0.0131 / (0.014 + 0.0131 + 0.0377) # =0.2
#Pop ID explains 20% of the variance.

library(effects)
plot(allEffects(mixed.lmer_r))
anova(mixed.lmer_r)
#library(broom)
#tidy(mixed.lmer_r, effects = "fixed")
#glance(mixed.lmer_r)
confint(mixed.lmer_r)
```

Multiple Regression

Best Subsets to determine the model with the highest r squared value. 
```{r}
rmod.fit <- regsubsets(YII ~ tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data=recovery.clim.ca.tl.2, nvmax=10)
sum.rmod <- summary(rmod.fit)
sum.rmod
plot(sum.rmod$adjr2,xlab = "Number of Variables", ylab = "Adjusted RSq")
adj_r2_max = which.max(sum.rmod$adjr2) #9
points(adj_r2_max, sum.rmod$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)
# Forward Stepwise Selection
regfit_fwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca.tl.2, nvmax = 10, method = "forward")
summary(regfit_fwd)
# Backward Stepwise Selection
regfit_bwd = regsubsets(YII~tmean + Elevation + tmax + tdmean + vpdmax + tmin + PPT + vpdmin + solclear + soltrans, data = recovery.clim.ca.tl.2, nvmax = 10, method = "backward")
summary(regfit_bwd)
```

Creating the multiple regression model
```{r}
rmod <- lm(YII ~ tmean + Elevation + tdmean + vpdmax + vpdmin + soltrans, data=recovery.clim.ca.tl.2)
summary(rmod) #adjusted r squared = 0.357
Anova(rmod, type=2)
par(mfrow=c(2,2))
plot(rmod)
par(mfrow=c(1,1))
```

Checking correlation of variables. 
```{r}
cor.table(recovery.clim.ca.tl.2[,c(21,23:27,29,31,32)])
rmod <- lm(YII ~ tmean)
cor_data <- cor((recovery.clim.ca[,c(21,23:27,29,31,32)]))
cor_data
cor_matrix_rm <- cor_data                # Modify correlation matrix
cor_matrix_rm[upper.tri(cor_matrix_rm)] <- 0
diag(cor_matrix_rm) <- 0
cor_matrix_rm
#To remove highly correlated variables
recovery.clim.ca.4 <- recovery.clim.ca[ , !apply(cor_matrix_rm,2,function(x) any(x > 0.99))]
```

Plotting YII against climatic variables
```{r}
p1 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Mean Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p2 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p3 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p4 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=Elevation, y=YII)) +
  geom_point() + theme_bw() +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p1, p2, p3, p4, ncol=2)
p5 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=PPT, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Precipitation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p6 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=tdmean, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Dew Point Temperature") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p7 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=vpdmax, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Max Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p8 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=vpdmin, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Min Annual Vapor Pressure Deficit") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p9 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=solclear, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Solar Radiation") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
p10 <- ggplot(data=recovery.clim.ca.tl.2, aes(x=soltrans, y=YII)) +
  geom_point() + theme_bw() + xlab("Average Annual Cloud Transmittance") +
  geom_smooth(method="lm", alpha=0.3, linetype="longdash", color="#0072B2")
grid.arrange(p5, p6, p7, p8, p9, p10, ncol=3)
```











